### 2017-09-17 coverletter.tex

Dear Editor,

This submitted article encompass the work that we have been doing and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.

EEG Processing and BCI are research areas in which we may be able to contribute as engineers and researchers in order to provide new methods and tools that could be used to improve the communication or treatment for persons affected by neuro-degenerative diseases.

We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its successful application to identify ERP markers within EEG signals from ALS patients is relevant for this special issue and that it will contribute with a novel tool to analyse EEG signals.

Yours truly and deeply greatfull for your help
Rodrigo Ramele
PhD Candidate
ITBA University 
Argentina